---
title: "CIA2"
author: "Abhilash M (1828601)"
date: "15 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown  
### __INTRODUCTION__
##### The dataset chosen is Car: Stated Preferences for Car Choice. This dataset consists of the choices of customers of a car and their preferred choice based on various other parameters.

### Description
##### a cross-section
##### number of observations : 4654
##### observation : individuals
##### country : United States

### Data Dictionary
#####A dataframe containing :

########1  choice
#####   choice of a vehicle among 6 propositions
######     1. Regular Car
######     2. SportUV
######     3. SportsCar
######     4. Stwagon
######     5. Truck
######     6. Van

#######2  college
#####   college education ?

#######3   hsg2
#####   size of household greater than 2 ?

#######4    coml5
#####   commute lower than 5 miles a day ?

#######5     typez
#####   body type, one of regcar (regular car), sportuv (sport utility vehicle), sportcar, stwagon (station wagon), truck, van, for each proposition z from 1 to 6

#######6     fuelz
#####   fuel for proposition z, one of gasoline, methanol, cng (compressed natural gas), electric.

#######7     pricez
#####   price of vehicle divided by the logarithm of income

#######8     rangez
#####   hundreds of miles vehicle can travel between refuelings/rechargings

#######9     accz
#####   acceleration, tens of seconds required to reach 30 mph from stop

#######10     speedz
#####   highest attainable speed in hundreds of mph

#######11     pollutionz
#####   tailpipe emissions as fraction of those for new gas vehicle

#######12     sizez
#####   0 for a mini, 1 for a subcompact, 2 for a compact and 3 for a mid-size or large vehicle

#######13     spacez
#####   fraction of luggage space in comparable new gas vehicle

#######14     costz
#####   cost per mile of travel (tens of cents) : home recharging for electric vehicle, station refueling otherwise

#######15     stationz
#####   fraction of stations that can refuel/recharge vehicle

### Analysis
```{r cars}
cars<-read.csv("C:/Users/Abhilash M Reddy/Desktop/SM/CIA 2/Car.csv")
summary(cars)
library(ggplot2)
library(grid)
library(gridBase)
library(gridExtra)
library(lattice)
library(vcd)
```
#### Objective 1
##### To identify the choice of car based on the repondents who donot have college education
```{r}
ggplot(cars, aes(x=choice, fill=college))+geom_bar(fill="tomato3")+labs(title = "Choice of car of people without college education")
```

##### The graph indicates that choice 5 is the highest chosen car among the respondents, followed by choice3 & choice1. Choice 2,4 and 6 are among the bottom 3 choices of the respondents.

```{r}

p1<-ggplot(cars, aes(choice, fill = type1))+   
  geom_bar(position = "fill")+   facet_wrap(~hsg2)

p2<-ggplot(cars, aes(choice, fill = type2))+   
  geom_bar(position = "fill")+   facet_wrap(~hsg2)

p3<-ggplot(cars, aes(choice, fill = type3))+   
  geom_bar(position = "fill")+   facet_wrap(~hsg2)

p4<-ggplot(cars, aes(choice, fill = type4))+   
  geom_bar(position = "fill")+   facet_wrap(~hsg2)

p5<-ggplot(cars, aes(choice, fill = type5))+   
  geom_bar(position = "fill")+   facet_wrap(~hsg2)

p6<-ggplot(cars, aes(choice, fill = type6))+   
  geom_bar(position = "fill")+   facet_wrap(~hsg2)

t<-textGrob("Choice of car vs body type among household size greater or lesser than 2")
```
#### Objective 2 
##### To graphically visualize the popularity of vehicle made on the body type among household size greater or lesser than 2 among various choices.
```{r}
grid.arrange(t,p1,p2)
```

#####From the graph it is seen that the choice of cars among both household size greater than or lesser than 2 based on body type 1 is regular car. However, in terms of the next two popular choices, trucks are more prominent than vans in household size less than 2, whereas trucks and vans have similar count among the various choices of cars chosen.

#####Similarly in body type2, the trucks have a major share compared to other cars in both scenarios of household size greater and lesser than 2. However, it is to note that the number of sports car and sports utility cars in this body type are very less or negligible indicating least popularity among the choices.
```{r}
grid.arrange(t,p3,p4)
```

#####In the type3 body type of cars, it is seen that regular cars are a more popular choice of cars among both household sizes lesser than or greater than 2. This is followed by trucks and vans as similar to body type1. However, it is to note that sports car and sports utility vehicle are less in household size less than 2 and sports cars are negligible in the body type3 among household size greater than 2. This indicates the unpopularity of sports car in this body type.

#####In type4 body type, vans have more popularity among various choices and also among the different household sizes. This is followed by station wagons and trucks with household size less than 2 having an inclination more towards trucks than station wagons and viceversa for household size greater than 2. Sports car and sports utility vehicle have least popularity among both types of households over various choices.

```{r}
grid.arrange(t,p5,p6)
```

#####In body type5, regular cars have the highest popularity among all the choices in both household categories. They are followed by trucks and van in which trucks play a more popular role compared to vans in household size less than 2 and viceversa in the household sizes greater than 2. Sports car and sports utility vehicle are a little more significant in this body type among household size less than 2 whereas sports car remains unpopular or least popular in the other segment.

#####In body type6, station wagons play more as popular car choice compared to other cars, followed by trucks and vans. Here trucks are more popular than vans in household size greater than 2 and viceversa in household size lesser than 2. Sports and sports utility vehicle are unpopular in this body type as well.

####Objective 3
#####To visually identify the choice made by the customer based on whether they commute less than or more than 5 miles a day.

```{r}
mosaicplot(~ coml5+  choice, data = cars, color = T)
```

#####Based on the above graph, it is that people who commute less than 5 miles a day perfer more among choice3 and choice5 models of car that is sports car and trucks respectively, whereas people commuting more than 5 miles a day prefer choice5 that is a truck over another choices. However, compared to people who commute more than 5 miles a day, the choice of sports car is more among the ones who commute less than 5 miles a day.

####Source
#####McFadden, Daniel and Kenneth Train (2000) "Mixed MNL models for discrete response", Journal of Applied Econometrics, 15(5), 447-470.

####References
#####Journal of Applied Econometrics data archive : http://qed.econ.queensu.ca/jae/.

####See Also
#####Index.Source, Index.Economics, Index.Econometrics, Index.Observations



